---
title: "GRIN Epilepsy Trio Analysis (Gemini)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    lib_dir: libs
css: styles.css
---
```{r libs, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, results='hide', fig.path='{{SLINK}}/'}
library(knitr)
library(DT)
library(plyr)
library(dplyr)
library(GenomicRanges)
options(digits = 2)
options(width=1200)
```

```{r functions, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, results='hide', fig.path='{{SLINK}}/'}
## http://www.r-bloggers.com/identifying-records-in-data-frame-a-that-are-not-contained-in-data-frame-b-%E2%80%93-a-comparison/
thisnotinthat <- function(x.1,x.2,...){
  x.1p <- do.call("paste", x.1)
  x.2p <- do.call("paste", x.2)
  x.1[! x.1p %in% x.2p, ]
}

viewDataTable <- function(dat){

  cols2hide <- grep('Description|Variants|Perc|chromosome|position|rsid|^ref$|^alt$|^gene$|
                    impact_severity|rvis_perc|depths|gt_quals|family_genotypes',
                    colnames(dat), invert = TRUE)

  DT::datatable(dat,
                extensions = c('Buttons'),
                selection = "single",
                filter = "bottom",
                options = list(
                  columnDefs = list(list(visible = FALSE, targets = cols2hide)),
                  dom = 'Bfrtip',
                  buttons = list('colvis','pageLength', 'copy','print',
                                 list(extend = "collection",
                                      buttons = c('csv', 'excel', 'pdf'),
                                      text = 'Download'
                                 )),
                  searchHighlight = TRUE,
                  lengthMenu = list(c(5, 10, 15, 20, 25, -1), c('5', '10', '15', '20', '25', 'All')),
                  initComplete = DT::JS("function(settings, json) {",
                                        "$(this.api().table().header()).css({'background-color': '#003366', 'color': '#fff'});",
                                        "}"),
                  scrollX = TRUE
                ))
}
```

## Filters
<blockquote>
We used the following filters to reduce the number of reported false positives:

1. Spurious inherited variants can often appear due to insufficient sequence coverage so the **mininum sequence depth** was set to **15**.

2. We are interested in mutations that are likely to have functional consequence so only variants with **Medium or High impact severity** were used.
</blockquote>

## De novo analysis
<blockquote>
Genotype Requirements:
  
1. all affecteds must be het
2. all unaffected must be homref or homalt
3. at least 1 affected kid must have unaffected parents
4. if an affected has affected parents, it’s not de_novo
5. all affected kids must have unaffected (or no) parents
6. warning if none of the affected samples have parents.
</blockquote>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Variant summary for de novo variants</h3>
</div>
<div class="panel-body">
This represents a brief summary of the de novo variants.

```{r denovo_summary, results='asis', warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}

if(length(denovo) == 1 && denovo == 0)
{
  varcount <- 0
  showChunk<-FALSE
} else {
  showChunk<-TRUE
  dt <- plyr::count(denovo , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = round(prop.table(dt$freq)*100,2)))
}
```

```{r denovo_pp, warnings = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,fig.path='{{SLINK}}/'}
gr1 <- with(denovo, GRanges(chrom, IRanges(start = start, end = end))) 
gr2 <- with(denovogear, GRanges(chr, IRanges(start = pos, end = pos)))
type1 <- findOverlaps(query = gr1, subject = gr2, type = 'any')
denovo$denovo_prob <- NA
denovo[queryHits(type1),'denovo_prob'] <- denovogear[subjectHits(type1),'pp_dnm']
denovo <- denovo[order(denovo$denovo_prob, decreasing = TRUE),]

denovo$start <- ifelse(denovo$end-denovo$start==1, denovo$end, paste0(denovo$start,'-',denovo$end))
colnames(denovo)[grep('chr|start|vcf_id', colnames(denovo))] <- c('chromosome','position','rsid')
denovo$end <- NULL
```

```{r denovo_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- length(unique(denovo$variant_id))
```

**There are `r varcount` de novo variants.**
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 1: De novo variants in recurrent genes</h3>
</div>
<div class="panel-body">
These are variants found in the recurrent genes associated with recurrent mutations in epilepsy.

```{r denovo_recurrent, results='asis', warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
denovo %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent 
varcount <- length(unique(in_recurrent$variant_id))
if(varcount>0){
viewDataTable(dat = in_recurrent)
}
```
**There are `r varcount` de novo variants in recurrent genes.**

</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 2: De novo variants in literature genes with low tolerance</h3>
</div>
<div class="panel-body">
These are the remaining variants cited in the literature genes cited in the existing literature. Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r denovo_literature, results='asis', warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
denovo %>% dplyr::filter(literature_genes != "None") %>% dplyr::filter(as.numeric(rvis_score)<=0) %>% thisnotinthat(in_recurrent) -> in_literature
varcount <- length(unique(in_literature$variant_id))
if(varcount){
viewDataTable(dat = in_literature)
}
```
**There are `r varcount` remaining variants found in literature genes.**

</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 3: De novo coding variants not in the above</h3>
</div>
<div class="panel-body">
These are remaining variants that are located in the coding sequence of at least one transcript.

```{r denovo_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
denovo %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
varcount <- length(unique(codingonly$variant_id))
if(varcount){
viewDataTable(dat = codingonly)
}
```
**There are `r varcount` remaining variants found in coding genes**
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Other de novo variants</h3>
</div>
<div class="panel-body">
These are the remaining de novo variants

```{r denovo_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(denovo, in_literature)
others <- thisnotinthat(others, codingonly)

varcount <- length(unique(others$variant_id))
if(varcount)
{
  viewDataTable(dat = others)
}
```
**There are `r varcount` remaining variants**
</div>
</div>


## Autosomal recessive homozygous
<blockquote>
Genotype Requirements:
  
1. all affecteds must be hom_alt
2. no unaffected can be hom_alt (can be unknown)
3. if parents exist they must be unaffected and het for all affected kids
4. if there are no affecteds that have a parent, a warning is issued.
</blockquote>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Variant summary for autosomal recessive homozygous variants</h3>
</div>
<div class="panel-body">
This represents a brief summary of the autosomal recessive homozygous variants.

```{r arhomo_summary, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}

if(length(arhomo) == 1 && arhomo == 0)
{
  varcount <- 0
  showChunk<-FALSE
} else {
  showChunk<-TRUE
  dt <- plyr::count(arhomo , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = round(prop.table(dt$freq)*100,2)))
}
```


```{r arhomo_pp, warnings = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,fig.path='{{SLINK}}/'}
arhomo$start <- ifelse(arhomo$end-arhomo$start==1, arhomo$end, paste0(arhomo$start,'-',arhomo$end))
colnames(arhomo)[grep('chr|start|vcf_id', colnames(arhomo))] <- c('chromosome','position','rsid')
arhomo$end <- NULL
```

```{r arhomo_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- length(unique(arhomo$variant_id))
```

There are `r varcount` autosomal recessive homozygous variants.
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 1: Autosomal recessive homozygous variants in recurrent genes</h3>
</div>
<div class="panel-body">
These are variants found in the genes associated with recurrent mutations in epilepsy.

```{r arhomo_recurrent, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
arhomo %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent
varcount <- length(unique(in_recurrent$variant_id))
if(varcount>0){
viewDataTable(dat = in_recurrent)
}
```

**There are `r varcount` de novo variants in recurrent genes.**
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 2: Autosomal recessive homozygous variants in literature genes with low tolerance</h3>
</div>
<div class="panel-body">

These are the remaining variants cited in the genes cited in the existing literature. Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r arhomo_literature, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
arhomo %>% dplyr::filter(literature_genes != "None") %>% dplyr::filter(as.numeric(rvis_score)<=0) %>% thisnotinthat(in_recurrent) -> in_literature
varcount <- length(unique(in_literature$variant_id))
if(varcount>0){
viewDataTable(dat = in_literature)
}
```
**There are `r varcount` remaining variants found in literature genes.**
</div>
</div>


<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 3: Autosomal recessive homozygous variants not in the above</h3>
</div>
<div class="panel-body">

These are remaining variants that are located in the coding sequence of at least one transcript.

```{r arhomo_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
arhomo %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
varcount <- length(unique(codingonly$variant_id))

if(varcount > 0){
viewDataTable(dat = codingonly)
}
```
**There are `r varcount` remaining variants found in coding genes**
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Other autosomal recessive homozygous variants</h3>
</div>
<div class="panel-body">

These are the remaining autosomal recessive homozygous variants

```{r arhomo_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(arhomo, in_literature)
others <- thisnotinthat(others, codingonly)

varcount <- length(unique(others$variant_id))
if(varcount)
{
  viewDataTable(dat = others)
}
```
**There are `r varcount` remaining variants**
</div>
</div>

## Compound Heterozygotes

Many autosomal recessive disorders are caused by compound heterozygotes. Unlike canonical recessive sites where the same recessive allele is inherited from both parents at the _same_ site in the gene, compound heterozygotes occur when the individual’s phenotype is caused by two heterozygous recessive alleles at _different_ sites in a particular gene.
<blockquote>
Genotype Requirements:
  
1. All affected individuals must be heterozygous at both sites.
2. No unaffected can be homozygous alterate at either site.
3. Neither parent of an affected sample can be homozygous reference at both sites.
4. If any unphased-unaffected is het at both sites, the site will be give lower priority
5. No phased-unaffected can be heterozygous at both sites.
6. Remove candidates where an affected from the same family does NOT share the same het pair.
7. Sites are automatically phased by transmission when parents are present in order to remove false positive candidates.
</blockquote>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Variant summary for Compound Heterozygotes</h3>
</div>
<div class="panel-body">

This represents a brief summary of the Compound Heterozygotes.

```{r cmpdhet_summary, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}

if(length(cmpdhet) == 1 && cmpdhet == 0)
{
  varcount <- 0
  showChunk<-FALSE
} else {
  showChunk<-TRUE
  dt <- plyr::count(cmpdhet , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = round(prop.table(dt$freq)*100,2)))
}
```

```{r cmpdhet_pp, warnings = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,fig.path='{{SLINK}}/'}
cmpdhet$start <- ifelse(cmpdhet$end-cmpdhet$start==1, cmpdhet$end, paste0(cmpdhet$start,'-',cmpdhet$end))
colnames(cmpdhet)[grep('chr|start|vcf_id', colnames(cmpdhet))] <- c('chromosome','position','rsid')
cmpdhet$end <- NULL
```

```{r cmpdhet_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- length(unique(arhomo$variant_id))
```

**There are `r varcount` compound heterozygotes.**
</div>
</div>

#### Triage comphet
<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 1: Compound Heterozygotes in recurrent genes</h3>
</div>
<div class="panel-body">
These are variants found in the genes associated with recurrent mutations in epilepsy.

```{r cmpdhet_recurrent, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
cmpdhet %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent
varcount <- length(unique(in_recurrent$variant_id))
if(varcount){
viewDataTable(dat = in_recurrent)
}
```
**There are `r varcount` remaining variants found in recurrent genes.**
</div>
</div>


<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 2: Compound Heterozygotes in literature genes with low tolerance</h3>
</div>
<div class="panel-body">
These are the remaining variants cited in the genes cited in the existing literature. 
Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r cmpdhet_literature, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
cmpdhet %>% dplyr::filter(literature_genes != "None") %>% dplyr::filter(as.numeric(rvis_score)<=0) %>% thisnotinthat(in_recurrent) -> in_literature
varcount <- length(unique(in_literature$variant_id))
if(varcount){
viewDataTable(dat = in_literature)
}
```
**There are `r varcount` remaining variants found in literature genes.**
</div>
</div>


<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 3: Comp het coding variants not in the above</h3>
</div>
<div class="panel-body">
These are remaining variants that are located in the coding sequence of at least one transcript.

```{r cmpdhet_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
cmpdhet %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
varcount <- length(unique(codingonly$variant_id))
if(varcount){
viewDataTable(dat = codingonly)
}
```
**There are `r varcount` remaining variants found in coding genes**
</div>
</div>



<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Other Compound Heterozygotes</div>
<div class="panel-body">
These are the remaining compound heterozygotes

```{r cmpdhet_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(cmpdhet, in_literature)
others <- thisnotinthat(others, codingonly)

varcount <- length(unique(others$variant_id))
if(varcount)
{
  viewDataTable(dat = others)
}
```
**There are `r varcount` remaining variants**
</div>
</div>
