---
title: "GRIN Epilepsy Trio Analysis (Gemini)"
author: "Jeremy Leipzig"
date: "`r Sys.Date()`"
output: html_document
css: styles.css
---
```{r libs, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, results='hide', fig.path='{{SLINK}}/'}
library(knitr)
library(DT)
library(plyr)
library(dplyr)
options(scipen = 999, stringsAsFactors = F, digits = 2)
```

```{r functions, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, results='hide', fig.path='{{SLINK}}/'}
# http://www.r-bloggers.com/identifying-records-in-data-frame-a-that-are-not-contained-in-data-frame-b-%E2%80%93-a-comparison/
thisnotinthat <- function(x.1,x.2,...){
  x.1p <- do.call("paste", x.1)
  x.2p <- do.call("paste", x.2)
  x.1[! x.1p %in% x.2p, ]
}

viewDataTable <- function(dat){
  DT::datatable(dat,
                extensions = c('Buttons'),
                selection = "single",
                filter = "bottom",
                options = list(
                  dom = 'Bfrtip',
                  buttons = list('colvis','pageLength', 'copy','print',
                                 list(extend = "collection",
                                      buttons = c('csv', 'excel', 'pdf'),
                                      text = 'Download'
                                 )),
                  searchHighlight = TRUE,
                  lengthMenu = list(c(5, 10, 15, 20, 25, -1), c('5', '10', '15', '20', '25', 'All')),
                  initComplete = JS("function(settings, json) {",
                                    "$(this.api().table().header()).css({'background-color': '#003366', 'color': '#fff'});",
                                    "}"),
                  scrollX = TRUE
                ))
}
```

# De novo analysis

Genotype Requirements:
  
1. all affecteds must be het
2. all unaffected must be homref or homalt
3. at least 1 affected kid must have unaffected parents
4. if an affected has affected parents, it’s not de_novo
5. all affected kids must have unaffected (or no) parents
6. warning if none of the affected samples have parents.

## Variant summary for de novo variants
This represents a brief summary of the de novo variants.

```{r denovo_summary, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
if(nrow(denovo) == 0)
{
  print("0 de novo variants found<br/>")
  showChunk<-FALSE
} else {
  showChunk<-TRUE
  dt <- plyr::count(denovo , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = prop.table(dt$freq)*100))
}
```

```{r denovo_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- unique(length(denovo$variant_id))
```

There are `r varcount` de novo variants.

#### Triage 1: De novo variants in recurrent genes

These are variants found in the recurrent genes associated with recurrent mutations in epilepsy.

```{r denovo_recurrent, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
in_recurrent <- denovo[which(denovo$recurrent_genes != "None"),]
# denovo %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent 
if (nrow(in_recurrent) == 0){
  print("No variants found in recurrent genes<br/>")
} else { 
  viewDataTable(dat = in_recurrent)
}
```

#### Triage 2: De novo variants in literature genes with low tolerance

These are the remaining variants cited in the literature genes cited in the existing literature. Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r denovo_literature, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
in_literature <- denovo[which(denovo$literature_genes != "None" & as.numeric(denovo$rvis_score) <= 0),]
in_literature <- thisnotinthat(x.1 = in_literature, x.2 = in_recurrent)
# denovo %>% dplyr::filter(literature_genes != "None") %>% 
#   dplyr::filter(rvis_score<=0) %>% 
#   thisnotinthat(in_recurrent) -> in_literature
if (nrow(in_literature) == 0){
  print("No remaining variants found in literature genes<br/>")
} else {
  viewDataTable(dat = in_literature)
}
```

#### Triage 3: De novo coding variants not in the above

These are remaining variants that are located in the coding sequence of at least one transcript.

```{r denovo_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
codingonly <- denovo[which(denovo$is_coding==1),]
codingonly <- thisnotinthat(codingonly, in_literature)
# denovo %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
if (nrow(codingonly) == 0){
  print("No remaining variants found in coding genes<br/>")
} else { 
  viewDataTable(dat = codingonly)
}
```

#### Other de novo variants

These are the remaining de novo variants

```{r denovo_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(denovo, codingonly)
# denovo %>% thisnotinthat(codingonly) -> others
viewDataTable(dat = others)
```

## Autosomal recessive homozygous

Genotype Requirements:
  
1. all affecteds must be hom_alt
2. no unaffected can be hom_alt (can be unknown)
3. if parents exist they must be unaffected and het for all affected kids
4. if there are no affecteds that have a parent, a warning is issued.

## Variant summary for autosomal recessive homozygous variants
This represents a brief summary of the autosomal recessive homozygous variants.

```{r arhomo_summary, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
if(nrow(arhomo) == 0)
{
  print("0 autosomal recessive variants found<br/>")
  showChunk<-FALSE
} else {
  showChunk<-TRUE
  dt <- plyr::count(arhomo , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = prop.table(dt$freq)*100))
}
```

```{r arhomo_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- unique(length(arhomo$variant_id))
```

There are `r varcount` autosomal recessive homozygous variants.

#### Triage 1: Autosomal recessive homozygous variants in recurrent genes

These are variants found in the genes associated with recurrent mutations in epilepsy.

```{r arhomo_recurrent, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
in_recurrent <- arhomo[which(arhomo$recurrent_genes != "None"),]
# arhomo %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent 
if (nrow(in_recurrent) == 0){
  print("No variants found in recurrent genes<br/>")
} else { 
  viewDataTable(dat = in_recurrent)
}
```

#### Triage 2: Autosomal recessive homozygous variants in literature genes with low tolerance

These are the remaining variants cited in the genes cited in the existing literature. Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r arhomo_literature, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
in_literature <- arhomo[which(arhomo$literature_genes != "None" & as.numeric(arhomo$rvis_score) <= 0),]
in_literature <- thisnotinthat(x.1 = in_literature, x.2 = in_recurrent)
#arhomo %>% dplyr::filter(literature_genes != "None") %>% dplyr::filter(rvis_score<=0) %>% thisnotinthat(in_recurrent) -> in_literature
if (nrow(in_literature) == 0){
  print("No remaining variants found in literature genes<br/>")
} else {
  viewDataTable(dat = in_literature)
}
```

#### Triage 3: Autosomal recessive homozygous variants not in the above

These are remaining variants that are located in the coding sequence of at least one transcript.

```{r arhomo_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
codingonly <- arhomo[which(arhomo$is_coding==1),]
codingonly <- thisnotinthat(codingonly, in_literature)
#arhomo %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
if (nrow(codingonly) == 0){
  print("No remaining variants found in coding genes<br/>")
} else { 
  viewDataTable(dat = codingonly)
}
```

#### Other autosomal recessive homozygous variants

These are the remaining autosomal recessive homozygous variants

```{r arhomo_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(arhomo, codingonly)
#arhomo %>% thisnotinthat(codingonly) -> others
viewDataTable(dat = others)
```

## Compound Heterozygotes

Many autosomal recessive disorders are caused by compound heterozygotes. Unlike canonical recessive sites where the same recessive allele is inherited from both parents at the _same_ site in the gene, compound heterozygotes occur when the individual’s phenotype is caused by two heterozygous recessive alleles at _different_ sites in a particular gene.

Genotype Requirements:
  
1. All affected individuals must be heterozygous at both sites.
2. No unaffected can be homozygous alterate at either site.
3. Neither parent of an affected sample can be homozygous reference at both sites.
4. If any unphased-unaffected is het at both sites, the site will be give lower priority
5. No phased-unaffected can be heterozygous at both sites.
6. Remove candidates where an affected from the same family does NOT share the same het pair.
7. Sites are automatically phased by transmission when parents are present in order to remove false positive candidates.

## Variant summary for Compound Heterozygotes

This represents a brief summary of the Compound Heterozygotes.

```{r cmpdhet_summary, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
if(nrow(cmpdhet) == 0)
{
  print("0 Compound Heterozygotes found<br/>")
  showChunk<-FALSE
} else {
  showChunk<-TRUE
  dt <- plyr::count(cmpdhet , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = prop.table(dt$freq)*100))
}
```

```{r cmpdhet_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- unique(length(arhomo$variant_id))
```

There are `r varcount` compound heterozygotes.

#### Triage 1: Compound Heterozygotes in recurrent genes

These are variants found in the genes associated with recurrent mutations in epilepsy.

```{r cmpdhet_recurrent, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
in_recurrent <- cmpdhet[which(cmpdhet$recurrent_genes != "None"),]
# cmpdhet %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent 
if (nrow(in_recurrent) == 0){
  print("No variants found in recurrent genes<br/>")
} else { 
  viewDataTable(dat = in_recurrent)
}
```

#### Triage 2: Compound Heterozygotes in literature genes with low tolerance

These are the remaining variants cited in the genes cited in the existing literature. 
Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r cmpdhet_literature, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
in_literature <- cmpdhet[which(cmpdhet$literature_genes != "None" & as.numeric(cmpdhet$rvis_score) <= 0),]
in_literature <- thisnotinthat(x.1 = in_literature, x.2 = in_recurrent)
# cmpdhet %>% dplyr::filter(literature_genes != "None") %>% dplyr::filter(rvis_score<=0) %>% thisnotinthat(in_recurrent) -> in_literature
if (nrow(in_literature) == 0){
  print("No remaining variants found in literature genes<br/>")
} else {
  viewDataTable(dat = in_literature)
}
```

#### Triage 3: Compound Heterozygotes not in the above

These are remaining variants that are located in the coding sequence of at least one transcript.

```{r cmpdhet_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
codingonly <- cmpdhet[which(cmpdhet$is_coding==1),]
codingonly <- thisnotinthat(codingonly, in_literature)
#cmpdhet %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
if (nrow(codingonly) == 0){
  print("No remaining variants found in coding genes<br/>")
} else { 
  viewDataTable(dat = codingonly)
}
```

#### Other Compound Heterozygotes

These are the remaining compound heterozygotes

```{r cmpdhet_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(cmpdhet, codingonly)
# cmpdhet %>% thisnotinthat(codingonly) -> others
viewDataTable(dat = others)
```

#### Done