---
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    lib_dir: libs
css: styles.css
params: 
    title: "GRIN Epilepsy Trio Analysis!"
---
---
title: `r params$title`
---
```{r libs, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, results='hide', fig.path='{{SLINK}}/'}
library(knitr)
library(DT)
library(plyr)
library(dplyr)
library(GenomicRanges)
options(digits = 2)
options(width=1200)
```

```{r functions, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, results='hide', fig.path='{{SLINK}}/'}
## http://www.r-bloggers.com/identifying-records-in-data-frame-a-that-are-not-contained-in-data-frame-b-%E2%80%93-a-comparison/
thisnotinthat <- function(x.1,x.2,...){
  x.1p <- do.call("paste", x.1)
  x.2p <- do.call("paste", x.2)
  x.1[! x.1p %in% x.2p, ]
}

viewDataTable <- function(dat){

  cols2hide <- grep('Description|Variants|Perc|chr|^pos$|rsid|^ref$|^alt$|^gene$|
                    impact_severity|rvis_score|depths|gt_quals|family_genotypes',
                    colnames(dat), invert = TRUE)

  DT::datatable(dat,
                extensions = c('Buttons'),
                selection = "single",
                filter = "bottom",
                options = list(
                  columnDefs = list(list(visible = FALSE, targets = cols2hide)),
                  dom = 'Bfrtip',
                  buttons = list('colvis','pageLength', 'copy','print',
                                 list(extend = "collection",
                                      buttons = c('csv', 'excel', 'pdf'),
                                      text = 'Download'
                                 )),
                  searchHighlight = TRUE,
                  lengthMenu = list(c(5, 10, 15, 20, 25, -1), c('5', '10', '15', '20', '25', 'All')),
                  initComplete = DT::JS("function(settings, json) {",
                                        "$(this.api().table().header()).css({'background-color': '#003366', 'color': '#fff'});",
                                        "}"),
                  scrollX = TRUE
                ))
}
```

## Filters
<blockquote>
We used the following filters to reduce the number of reported false positives:

1. Spurious inherited variants can often appear due to insufficient sequence coverage so the **mininum sequence depth** was set to **15**.

2. We are interested in mutations that are likely to have functional consequence so only variants with **Medium or High impact severity** were used.
</blockquote>


## Autosomal recessive homozygous
<blockquote>
Genotype Requirements:
  
1. all affecteds must be hom_alt
2. no unaffected can be hom_alt (can be unknown)
3. if parents exist they must be unaffected and het for all affected kids
4. if there are no affecteds that have a parent, a warning is issued.
</blockquote>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Variant summary for autosomal recessive homozygous variants</h3>
</div>
<div class="panel-body">
This represents a brief summary of the autosomal recessive homozygous variants.

```{r arhomo_summary, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}

if(length(arhomo) == 1 && arhomo == 0)
{
  varcount <- 0
  showChunk <- FALSE
} else {
  showChunk <- TRUE
  dt <- plyr::count(arhomo , 'impact_so')
  viewDataTable(dat = data.frame(Description = dt$impact_so, 
                                 Variants = dt$freq, 
                                 Perc = round(prop.table(dt$freq)*100,2)))
}
```

```{r arhomo_rvis, warnings = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,fig.path='{{SLINK}}/'}
arhomo <- merge(arhomo, rvis, by.x = "gene", by.y = "GENE", all.x = T, sort = F)
```

```{r arhomo_pp, warnings = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,fig.path='{{SLINK}}/'}
arhomo$start <- ifelse(arhomo$end-arhomo$start==1, arhomo$end, paste0(arhomo$start,'-',arhomo$end))
colnames(arhomo)[grep('chr|start|vcf_id', colnames(arhomo))] <- c('chr','pos','rsid')
arhomo$end <- NULL
arhomo <- arhomo[,c(2:16,1,17:ncol(arhomo))]
```

```{r arhomo_filtered_variants, warning = FALSE, eval=showChunk, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
varcount <- length(unique(arhomo$variant_id))
```

There are `r varcount` autosomal recessive homozygous variants.
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 1: Autosomal recessive homozygous variants in recurrent genes</h3>
</div>
<div class="panel-body">
These are variants found in the genes associated with recurrent mutations in epilepsy.

```{r arhomo_recurrent, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
arhomo %>% dplyr::filter(recurrent_genes != "None") -> in_recurrent
varcount <- length(unique(in_recurrent$variant_id))
if(varcount>0)
{
  viewDataTable(dat = in_recurrent)
}
```

**There are `r varcount` de novo variants in recurrent genes.**
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 2: Autosomal recessive homozygous variants in literature genes with low tolerance</h3>
</div>
<div class="panel-body">

These are the remaining variants cited in the genes cited in the existing literature. Many of the literature-cited genes are likely tolerant, so this list is further filtered to only genes with an RVIS "ALL_0.1." value of <= 0.

```{r arhomo_literature, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
arhomo %>% dplyr::filter(literature_genes != "None") %>% dplyr::filter(as.numeric(rvis_score)<=0) %>% thisnotinthat(in_recurrent) -> in_literature
varcount <- length(unique(in_literature$variant_id))
if(varcount)
{
  viewDataTable(dat = in_literature)
}
```
**There are `r varcount` remaining variants found in literature genes.**
</div>
</div>


<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Triage 3: Autosomal recessive homozygous variants not in the above</h3>
</div>
<div class="panel-body">

These are remaining variants that are located in the coding sequence of at least one transcript.

```{r arhomo_coding, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
arhomo %>% dplyr::filter(is_coding==1) %>% thisnotinthat(in_literature) -> codingonly
varcount <- length(unique(codingonly$variant_id))
if(varcount)
{
  viewDataTable(dat = codingonly)
}
```
**There are `r varcount` remaining variants found in coding genes**
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<h3 class="panel-title">Other autosomal recessive homozygous variants</h3>
</div>
<div class="panel-body">

These are the remaining autosomal recessive homozygous variants

```{r arhomo_others, eval=showChunk, warning = FALSE, echo=FALSE, message=FALSE, cache=FALSE,  fig.path='{{SLINK}}/'}
others <- thisnotinthat(arhomo, in_literature)
others <- thisnotinthat(others, codingonly)

varcount <- length(unique(others$variant_id))
if(varcount)
{
  viewDataTable(dat = others)
}
```
**There are `r varcount` remaining variants**
</div>
</div>
